The necessity of hypermedia RDF and an approach to achieve it

### 2012-04-13 eswc-2012/workshop2.tex
redo HATEOAS discussion
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@229 43ef1190-90fd-4dd7-a769-d0121f02b80e

In a read-only situation, the HATEOAS constraint requires that the
application can navigate from one resources to another using the
hypermedia links in the present resource \emph{only}, they should not
require any out-of-band information. It is very common that
specifications that advertise themselves as RESTful has a read-write
API that is specified in terms of URI structure as well as HTTP
verbs. This is not a violation of the HATEOAS constraint \textit{per
  se}, but it would usually be superfluous as the same information
\emph{must} be available in a hypermedia message in order to satisfy
the HATEOAS constraint. The constraint does not require this
information to be available in \emph{all} messages, nor does it
require that all resources can be reached from any other resource, and
so, it is not a very strict constraint.

Since RDF is built on URIs, many of which can be dereferenced to
obtain further links, it lends itself well to RESTful protocols, but
it is debatable whether a protocol can be said to be RESTful if the
links don't enable any useful interactions. Importantly, if there is
to be such a thing as a RESTful read-write Semantic Web protocol,
there must be something in the RDF itself that can be used by
practical applications to write data. Therefore, whether the HATEOAS
constraint is satisfied must be judged on the basis of the practical
applications the protocol enables.

>>>>

In a read-only situation, the HATEOAS constraint requires that the
application can navigate from one resources to another using the
hypermedia links in the present resource \emph{only}, they should not
require any out-of-band information. Since RDF is built on URIs, many
of which can be dereferenced to obtain further links, it will
therefore almost always satisfy the HATEOAS constraint in the
read-only case.

Contrast the read-write situation: It is very common that
specifications that advertise themselves as RESTful has an 
API that is specified in terms of URI structure as well as HTTP
verbs. This is not a violation of the HATEOAS constraint \textit{per
  se}, but it would usually be superfluous as the same information
\emph{must} be available in a hypermedia message in order to satisfy
the HATEOAS constraint. The constraint does not require this
information to be available in \emph{all} messages, and
so, it is not a very strict constraint.

It is debatable whether a protocol can be said to be RESTful if the
links don't enable any useful interactions. Even in the read-only
case, we must carefully add links that enable the desired interactions
to take place, e.g. linking a SPARQL endpoint. Importantly, if there is
to be such a thing as a RESTful read-write Semantic Web protocol,
there must be something in the RDF itself that can be used by
practical applications to write data. Therefore, whether the HATEOAS
constraint is satisfied must be judged on the basis of the practical
applications the protocol enables.
